\documentclass[letterpaper,12pt]{article}
\setlength{\headheight}{0ex}
\setlength{\headsep}{0ex}
\input{hogg_nasa}
\pagestyle{empty}
\begin{document}

\paragraph{David W. Hogg} http://cosmo.nyu.edu/hogg/

\paragraph{Professional Preparation:}~\\[3pt]
\begin{tabular}{llll}
\textsl{postdoctoral}
  & Institute for Advanced Study          & Astrophysics & 1997--2001\\
\textsl{graduate}
  & California Institute of Technology    & Physics & PhD 1998\\
\textsl{undergraduate}
  & Massachusetts Institute of Technology & Physics & SB 1992\\
\end{tabular}
           
\paragraph{Appointments:}~\\[3pt]
\begin{tabular}{lll}
Professor           & Department of Physics, New York University & 2014--present \\
Professor           & Center for Data Science, New York University & 2015--present \\
Senior Staff Scientist & Max-Planck-Institut f\"ur Astronomie    & 2012--present \\
Associate Professor & Department of Physics, New York University & 2007--2014 \\
Assistant Professor & Department of Physics, New York University & 2001--2007 \\
Long-term member    & Astrophysics, Institute for Advanced Study & 1997--2001 \\
\end{tabular}

\paragraph{Administrative Roles:}~\\[3pt]
\begin{tabular}{ll}
Deputy Director, NYU Center for Data Science & 2014--2015 \\
Executive Director, Moore--Sloan Data Science Environment at NYU & 2014--2015 \\
Director of Undergraduate Studies, NYU Department of Physics & 2008--2015 \\
\end{tabular}

\paragraph{Five Recent, Relevant Publications:}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Sch\"olkopf,~B., Hogg,~D.~W., Wang,~D., Foreman-Mackey,~D., Janzing,~D., Simon-Gabriel,~C.-J., \& Peters,~J., 2016,
{Modeling confounding by half-sibling regression},
\textit{Proc.\,Natl.\,Acad.\,Sci.}\ \textbf{113}(27) 7391–-7398.
\item
Wang,~D., Hogg,~D.~W., Foreman-Mackey,~D., \& Sch\"olkopf,~B., 2016,
{A Causal, Data-driven Approach to Modeling the Kepler Data},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{128} 094503.
\item
Ness,~M., Hogg,~D.~W., Rix,~H.-W., Ho,~A.~Y.~Q., \& Zasowski,~G., 2015,
{\project{The~Cannon}: A data-driven
approach to stellar label determination},
\textit{Astrophys.\,J.}\ \textbf{808} 16.
\item
Ambikasaran,~S., Foreman-Mackey,~D., Greengard,~L., Hogg,~D.~W., \& O'Neil,~M., 2015,
{Fast direct methods for Gaussian Processes and the analysis of \textsc{nasa} \project{Kepler} Mission Data},
\textit{IEEE Transactions on Pattern Analysis and Machine Intelligence}\ \textbf{PP} (99) 1,1.
\item
Foreman-Mackey~D., Hogg,~D.~W., \& Morton,~T.~D., 2014,
{Exoplanet population inference and the abundance of Earth analogs from noisy, incomplete catalogs},
\textit{Astrophys.\,J.}\ \textbf{795} 64.
\end{itemize}

\paragraph{Five Other Publications:}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Fergus,~R., Hogg,~D.~W., Oppenheimer,~R., Brenner,~D. \& Pueyo,~L., 2014,
{\project{S4}: A spatial-spectral model for speckle suppression},
\textit{Astrophys.\,J.}\ \textbf{794} 161.
\item
Foreman-Mackey,~D., Hogg,~D.~W., Lang,~D., \& Goodman,~J., 2013,
{\project{emcee}:\ The MCMC Hammer},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{125} 306--312.
\item
Bovy,~J. \etal, 2012,
{Photometric redshifts and quasar probabilities from a single, data-driven generative model},
\textit{Astrophys.\,J.}\ \textbf{749} 41.
\item
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011,
{Extreme deconvolution:\ Inferring complete distribution functions from noisy, heterogeneous and incomplete observations},
\textit{Ann.\,Appl.\,Stat.}\ \textbf{5} 1657--1677.
\item
Eisenstein,~D.~J., \etal, 2005,
{Detection of the baryon acoustic peak in the large-scale correlation function of \project{Sloan Digital Sky Survey} Luminous Red Galaxies},
\textit{Astrophys.\,J.}\ \textbf{633} 560--574.
\end{itemize}

\paragraph{Synergistic Activities:}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Co-author of open-source code
for performing tuned Monte Carlo sampling in inference contexts (http://dan.iel.fm/emcee/),
for fast Gaussian process regression and marginalization with enormous matrices (http://dan.iel.fm/george/),
for estimating densities given heterogeneous observations (http://code.google.com/p/extreme-deconvolution/),
for performing classification and target selection (http://code.google.com/p/xdtarget/),
and
for calibration of astronomical images (http://Astrometry.net/).
\item
Author of pedagogical texts on arXiv.org for cosmology, fundamental
measurement in astronomy, and---recently---statistical inference.
These texts are associated with publicly released code.
\item
Organizer of Hack Days at the American Astronomical Society winter meetings (2013--present).
Co-organizer of annual Astro Hack Week (2014--present).
\item
Recipient of various teaching awards: 
NYU ``Golden Dozen'' Teaching Award (2004);
Princeton University Engineering Council Teaching Award (2000);
Caltech Undergraduate Teaching Award (1996).
Gives occasional public lectures to increase public understanding of
science, most recently in downtown Abu Dhabi (audience of 250) and at
the NYC ``Secret Science Club'' (audience of 600).
\item
Participant in committees in service to the astronomical community,
recently including
White House OSTP Astronomy and Astrophysics Advisory Committee (2014--present),
\Spitzer\ Science Center Oversight Committee, (2008--present),
\project{Kepler} Exoplanet Occurence Rate review (2014),
\project{NASA Extragalactic Database} Users Committee (2006--2013),
and
Panel Chair, \Spitzer\ Time Allocation Committee (2005).
\end{itemize}

\paragraph{Graduate Advisor and Postdoctoral Sponsor:}
$\bullet$~Roger Blandford (Caltech PhD advisor; now at Stanford);
$\bullet$~John Bahcall (IAS postdoc sponsor; now deceased).

\paragraph{Graduate Students Advised:}
$\bullet$~Jo Bovy (PhD 2011);
$\bullet$~Daniel Foreman-Mackey (PhD 2015);
$\bullet$~Fengji Hou (PhD 2014);
$\bullet$~Tao Jiang (PhD 2012);
$\bullet$~Dustin Lang (PhD 2009; co-advised by Sam Roweis, Toronto);
$\bullet$~Alex Malz (current);
$\bullet$~Morad Masjedi (PhD 2007);
$\bullet$~Mohammadjavad Vakili (current);
$\bullet$~Dun Wang (current);
$\bullet$~Ronin RongYing Wu (PhD 2011);
$\bullet$~Adi Zolotov (PhD 2011; co-advised by Beth Willman, Haverford).

\paragraph{Postdoctoral Scholars Advised:}
$\bullet$~Andreas Berlind (now at Vanderbilt);
$\bullet$~Michael Blanton (now at NYU);
$\bullet$~Ross Fadely (now at Insight Data Science);
$\bullet$~Daniela Huppenkothen (current);
$\bullet$~Boris Leistedt (current);
$\bullet$~Gabriel Perez-Giz (now at the AAAS);
$\bullet$~Erin Sheldon (now at Brookhaven);
$\bullet$~Beth Willman (now at Haverford).

\end{document}
