\documentclass[12pt,letterpaper]{article}
\usepackage{fancyhdr, afterpage}
\usepackage{color, hyperref}
\usepackage{ifthen}

% put in the date -- or not
  \renewcommand{\today}{}

% hypertex insanity
  \definecolor{linkcolor}{rgb}{0,0,0.33}
  \definecolor{grey}{rgb}{0.5,0.5,0.5}
  \hypersetup{
    colorlinks=true,        % false: boxed links; true: colored links
    linkcolor=grey,        % color of internal links
    citecolor=linkcolor,    % color of links to bibliography
    filecolor=black,        % color of file links
    urlcolor=linkcolor      % color of external links
  }

% repeated text formatting
  \newcommand{\latin}[1]{\textsl{#1}}
  \newcommand{\etal}{\latin{et~al.}}
  \newcommand{\satellite}[1]{\textsl{#1}}
  \newcommand{\project}[1]{\textsl{#1}}
  \newcommand{\grantnumber}[2]{{\footnotesize{#1}}; #2,~\textsc{pi}}
% literature links--use doi if you can
  \newcommand{\doi}[2]{\href{http://dx.doi.org/#1}{{#2}}}
  \newcommand{\ads}[2]{\href{http://adsabs.harvard.edu/abs/#1}{{#2}}}
  \newcommand{\isbn}[1]{{\footnotesize(\textsc{isbn:}{#1})}}
  \newcommand{\arxiv}[2]{\href{http://arxiv.org/abs/#1}{{#2}\ (\texttt{arXiv:#1})}}
% define de-emphasize commands
  \newcommand{\deemph}[1]{\textcolor{grey}{\footnotesize{#1}}}
  \newcommand{\pubnumber}[1]{\deemph{{#1}.}}
% define heading command SERIOUS HACKING HERE
  \newcommand{\hoggheading}[1]{\addvspace{0.5ex}\pagebreak[2]%
    \par\textbf{#1}%
    \afterpage{\markboth{}{}}
    \nopagebreak}
% custom list characteristics
  \newcounter{refpubnum}
  \newcommand{\hogglist}{%
    \rightmargin=0in
    \leftmargin=0.18in
    \topsep=0ex
    \partopsep=0pt
    \itemsep=0.1ex
    \parsep=0pt
    \itemindent=-1.0\leftmargin
    \listparindent=0.0\leftmargin
    \settowidth{\labelsep}{~}
    \usecounter{refpubnum}
  }
  \newcommand{\note}[1]{{\small\par{#1}}}
% Margins and spaces
  \raggedright
  \setlength{\oddsidemargin}{0in}
  \setlength{\topmargin}{0in}
  \setlength{\headsep}{0.20in}
  \setlength{\headheight}{0.25in}
  \setlength{\textheight}{9.00in}
%  \addtolength{\textheight}{-\headsep}
%  \addtolength{\textheight}{-\headheight}
  \addtolength{\topmargin}{-\headsep}
  \addtolength{\topmargin}{-\headheight}
  \setlength{\textwidth}{6.50in}
  \setlength{\parindent}{0in}
  \setlength{\parskip}{1ex}
% Headings and footing
  \renewcommand{\headrulewidth}{0pt}
  \pagestyle{fancy}
  \lhead{\deemph{\leftmark}}
  \rhead{\deemph{\thepage}}
  \cfoot{}

\begin{document}\thispagestyle{empty}\sloppy\sloppypar\raggedbottom\frenchspacing
\textbf{\Large David W. Hogg} \hfill \url{http://cosmo.nyu.edu/hogg/}

\hoggheading{Professional Preparation}
\begin{list}{}{\hogglist}
\item
\deemph{undergraduate:} Massachusetts Institute of Technology, Physics, SB 1992
\item
\deemph{graduate:} California Institute of Technology, Physics, PhD 1998
\item
\deemph{postdoctoral:} Institute for Advanced Study, Astrophysics, 1997--2001
\end{list}

\hoggheading{Appointments}
\begin{list}{}{\hogglist}
\item
\deemph{current:} (Full) Professor of Physics and Data Science, New York University, 2014--.
\item
\deemph{current:} Group Leader, Astronomical Data Group, Center for Computational Astrophysics,
Flatiron Institute, New York, 2017--.
\item
\deemph{current:} Adjunct Senior Staff Scientist, Max-Planck-Institut f\"ur Astronomie,
Heidelberg, Germany, 2012--.
\item
Consultant, Flatiron Institute, New York, 2016--2017.
\item
Consultant, Simons Center for Data Analysis,
Simons Foundation, New York, 2015--2016.
\item
Associate Professor of Physics with tenure, New York University, 2007--2014.
\item
Visiting Scientist, Max-Planck-Institut f\"ur Astronomie,
Heidelberg, Germany, 2006--2011.
\item
Visiting Professor, Department of Astronomy and Astrophysics, Columbia
University, 2008
\item
Consultant, Google Inc., 2008.
\item
Assistant Professor of Physics, New York University, 2001--2007.
\item
Scholar in Residence, \satellite{Spitzer} Science Center, California
Institute of Technology, 2006
\item
Visiting Professor, Department of Physics, Massachusetts Institute of
Technology, 2005
\item
Lecturer (part-time), Department of Physics, Princeton University,
1998--2001.
\end{list}

\hypersetup{linkcolor=black}%
\hoggheading{Products (closely related)}%% use \etal when there are >9 authors!
\hypersetup{linkcolor=grey}%
\begin{list}{}{\hogglist}
\item
Masjedi,~M., Hogg,~D.~W., \& Blanton,~M.~R., 2008,
\doi{10.1086/586696}{The growth of luminous red galaxies by merging},
\textit{Astrophys.\,J.}\ \textbf{679} 260--268.
\item
Lang,~D., Hogg,~D.~W., Mierle,~K., Blanton,~M., \& Roweis,~S., 2010,
\doi{10.1088/0004-6256/139/5/1782}{\project{Astrometry.net}:\ Blind astrometric calibration of arbitrary astronomical images},
\textit{Astron.\,J.}\ \textbf{139} 1782--1800.
\item
Foreman-Mackey,~D., Hogg,~D.~W., Lang,~D., \& Goodman,~J., 2013,
\doi{10.1086/670067}{\project{emcee}:\ The MCMC Hammer},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{125} 306--312.
\item
Wang,~D., Hogg,~D.~W., Foreman-Mackey,~D., \& Sch\"olkopf,~B., 2016,
\doi{10.1088/1538-3873/128/967/094503}{A Causal, Data-driven Approach to Modeling the \satellite{Kepler} Data},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{128} 094503.
\item
Hahn,~C., Vakili,~M., Walsh,~K., Hearin,~A.~P., Hogg,~D.~W., \& Campbell,~D., 2017,
\doi{10.1093/mnras/stx894}{Approximate Bayesian computation in large-scale structure: constraining the galaxy-halo connection},
\textit{Mon.\,Not.\,R.\,Astr.\,Soc.}\ \textbf{469} 2791--2805.
\end{list}

\hypersetup{linkcolor=black}%
\hoggheading{Products (other significant)}%% use \etal when there are >9 authors!
\hypersetup{linkcolor=grey}%
\begin{list}{}{\hogglist}
\item\label{pub:Eisenstein2005}
Eisenstein,~D.~J., Zehavi,~I., Hogg,~D.~W., \etal, 2005,
\doi{10.1086/466512}{Detection of the baryon acoustic peak in the large-scale correlation function of \project{Sloan Digital Sky Survey} Luminous Red Galaxies},
\textit{Astrophys.\,J.}\ \textbf{633} 560--574.
\item
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011,
\doi{10.1214/10-AOAS439}{Extreme deconvolution:\ Inferring complete distribution functions from noisy, heterogeneous and incomplete observations},
\textit{Ann.\,Appl.\,Stat.}\ \textbf{5} 1657--1677.
\item
Oppenheimer,~B.~R. \etal, 2013,
\doi{10.1088/0004-637X/768/1/24}{Reconnaissance of the HR~8799 exosolar system.\ I.\ Near-infrared spectroscopy},
\textit{Astrophys.\,J.}\ \textbf{768} 24.
\item
Ness,~M., Hogg,~D.~W., Rix,~H.-W., Ho,~A.~Y.~Q., \& Zasowski,~G., 2015,
\doi{10.1088/0004-637X/808/1/16}{\project{The~Cannon}: A data-driven
approach to stellar label determination},
\textit{Astrophys.\,J.}\ \textbf{808} 16.
\item
El-Badry,~K., Ting,~Y.-S., Rix,~H.-W., Quataert,~E., Weisz,~D.~R., Cargile,~P., Conroy,~C., Hogg,~D.~W., Bergemann,~M., \& Liu,~C., 2018,
\doi{10.1093/mnras/sty240}{Discovery and Characterization of 3000+ Main-Sequence Binaries from \project{APOGEE} Spectra},
\textit{Mon.\,Not.\,R.\,Astr.\,Soc.}\ \textbf{476} 528--553.
\end{list}

\hoggheading{Synergistic activities}
\begin{list}{}{\hogglist}
\item
\satellite{Spitzer} Science Center Oversight Committee
(to advise the \satellite{Spitzer Space Telescope} project),
2008--present. Also served on the \project{\textsc{nasa} Extragalactic Database} Users Committee,
2006--2013.
\item
Involved in the every phase of the
\project{Sloan Digital Sky Survey \textsc{iv}} projects. Member of the
\project{SDSS-IV} Collaboration Council,
2013--present. Currently designing and fundraising \project{SDSS-V}.
\item
Astronomy and Astrophysics Advisory Committee
(established by the US Congress to advise \textsc{nsf}, \textsc{nasa}, and \textsc{doe} on areas of mutual interest)
2014--2017.
\item
Served in administrative roles at NYU, including
Director of Undergraduate Studies, Department of Physics, New York University, 2008--2015; and
Executive Director of the \textsc{usd}12.6M \project{Moore--Sloan Data Science Environment at \textsc{nyu}}, 2013--2015.
\item
Supervised the production of multiple open-source software projects, including \project{Astrometry.net}
and the \project{emcee} MCMC sampler.
\end{list}

\end{document}
