% this is a LaTeX file
%----------------------------------------------------------------------
% name:
%   hogg_cv.tex
% purpose:
%   Hogg's CV and list of publications
% to-do:
%   - convert all \ads{} to \doi{} -- the DOIs should exist now.
%   - spot-check links; these were all upgraded on 2012-11-07
%   - Is the ``what constrains'' paper published anywhere?
% comments:
%   - Don't list full author lists of 10 or more.
%   - If you want to include/remove the date, see \renewcommand{\today}.
%   - Swap 1/0 flag to include invited talks.
%----------------------------------------------------------------------

\documentclass[12pt,letterpaper]{article}
\usepackage{color}
\usepackage{fancyheadings}
\usepackage{hyperref}
\usepackage{ifthen}

% put in the date -- or not
  \renewcommand{\today}{2014 February 20}

% hypertex insanity
  \definecolor{linkcolor}{rgb}{0,0,0.4}
  \hypersetup{
    colorlinks=true,        % false: boxed links; true: colored links
    linkcolor=linkcolor,    % color of internal links
    citecolor=linkcolor,    % color of links to bibliography
    filecolor=linkcolor,    % color of file links
    urlcolor=linkcolor      % color of external links
  }

% repeated text formatting
  \newcommand{\latin}[1]{\textit{#1}}
  \newcommand{\etal}{\latin{et~al}}
  \newcommand{\satellite}[1]{\textsl{#1}}
  \newcommand{\project}[1]{\textsl{#1}}
  \newcommand{\usd}[1]{{#1}\,\textsc{usd}}
  \newcommand{\eur}[1]{{#1}\,\textsc{eur}}
% literature links--use doi if you can
  \newcommand{\doi}[2]{\href{http://dx.doi.org/#1}{{#2}}}
  \newcommand{\ads}[2]{\href{http://adsabs.harvard.edu/abs/#1}{{#2}}}
  \newcommand{\isbn}[1]{{\footnotesize(\textsc{isbn:}{#1})}}
  \newcommand{\arxiv}[2]{\href{http://arxiv.org/abs/#1}{{#2}}\ (\texttt{arXiv:#1})}
% define de-emphasize commands
  \definecolor{grey}{rgb}{0.5,0.5,0.5}
  \newcommand{\deemph}[1]{\textcolor{grey}{\footnotesize{#1}}}
  \newcommand{\pubnumber}[1]{\deemph{{#1}.}}
% define heading command
  \newcommand{\hoggheading}[1]{\addvspace{1ex}\pagebreak[2]%
    \par\textbf{#1}\nopagebreak}
% custom list characteristics
  \newcounter{refpubnum}
  \newcommand{\hogglist}{%
    \rightmargin=0in
    \leftmargin=0.15in
    \topsep=0ex
    \partopsep=0pt
    \itemsep=0.2ex
    \parsep=0pt
    \itemindent=-1.0\leftmargin
    \listparindent=0.0\leftmargin
    \settowidth{\labelsep}{~}
    \usecounter{refpubnum}
  }
  \newcommand{\note}[1]{{\small\par{#1}}}
% Margins and spaces
  \raggedright
  \setlength{\oddsidemargin}{0in}
  \setlength{\topmargin}{0in}
  \setlength{\headsep}{0.20in}
  \setlength{\headheight}{0.25in}
  \setlength{\textheight}{9.00in}
%  \addtolength{\textheight}{-\headsep}
%  \addtolength{\textheight}{-\headheight}
  \addtolength{\topmargin}{-\headsep}
  \addtolength{\topmargin}{-\headheight}
  \setlength{\textwidth}{6.50in}
  \setlength{\parindent}{0in}
  \setlength{\parskip}{1ex}
% Headings and footing
  \renewcommand{\headrulewidth}{0pt}
  \pagestyle{fancy}
  \lhead{\deemph{David W. Hogg}}
  \chead{\deemph{applicant biography}}
  \rhead{\deemph{\thepage}}
  \cfoot{}

\begin{document}\thispagestyle{empty}\sloppy\sloppypar\raggedbottom
\textbf{\Large David W. Hogg} \hfill \url{http://cosmo.nyu.edu/hogg/} \\[0.5ex]
Associate Professor, \\
Center for Cosmology and Particle Physics, Department of Physics, New York University \\[0.5ex]
Adjunct Senior Staff Scientist, \\
Max-Planck-Institut f\"ur Astronomie, Heidelberg, Germany

\hoggheading{Professional preparation}
\begin{list}{}{\hogglist}
\item
PhD 1998, Department of Physics, California Institute of Technology.
\item
SB 1992 (Major in Physics), Massachusetts Institute of Technology.
\end{list}

\hoggheading{Appointments}
\begin{list}{}{\hogglist}
\item
Associate Professor with tenure, Department of Physics, New York
University, 2007--present.
\item
Assistant Professor, Department of Physics, New York University,
2001--2007.
\item
Long-term member, Institute for Advanced Study, 1997--2001.
\end{list}

\hoggheading{Recent synergistic activities}
% reverse chronological by END date.
\begin{list}{}{\hogglist}
\item
Administrative roles:
Executive Director, \project{Moore--Sloan Data Science Environment at NYU}, 2014--present;
Deputy Director, Center for Data Science, New York University, 2014--present;
Director of Undergraduate Studies, Department of Physics, New York University, 2008--present.
\item
Community service:
\project{Sloan Digital Sky Survey IV} Collaboration Council,
2013--present;
\satellite{Spitzer} Science Center Oversight Committee,
2008--present;
\project{NASA Extragalactic Database} Users Committee,
2006--2013.
\item
Outreach:
Public talks about astronomy and data at the
  NYU Abu Dhabi Institute,
  University of Toledo,
  New York City Secret Science Club,
  and on YouTube for the Kavli Institute for Particle Astrophysics and Cosmology.
\item
Education:
Invited lecturer at workshops on data analysis at AISTATS and at Heidelberg IMPRS;
Co-organizer of a SAMSI workshop bringing together statisticians and astronomers to work on exoplanet data;
Creator of a few YouTube videos about probabilistic modeling of data;
Originator and co-organizer (with Kelle Cruz) of AAS Hack Days at the winter AAS meetings.
\end{list}

\hoggheading{Papers and projects highly relevant to the present proposal}
\begin{list}{}{\hogglist}
\item
Padmanabhan,~N., \etal, 2008,
\doi{10.1086/524677}{An improved photometric calibration of the \project{Sloan Digital Sky Survey} imaging data},
\textit{Astrophys.\,J.}\ \textbf{674} 1217--1233.
\item
Barron,~J.~T., Hogg,~D.~W., Lang,~D., \& Roweis,~S., 2008,
\doi{10.1088/0004-6256/136/4/1490}{Blind Date:\ Using proper motions to determine the ages of historical images},
\textit{Astron.\,J.}\ \textbf{136} 1490--1501.
\item
Lang,~D., Hogg,~D.~W., Jester,~S., \& Rix,~H.-W., 2009,
\doi{10.1088/0004-6256/137/5/4400}{Measuring the undetectable:\ Proper motions and parallaxes of very faint sources},
\textit{Astron.\,J.}\ \textbf{137} 4400--4411.
\item
Price-Whelan,~A.~M. \& Hogg,~D.~W., 2010,
\doi{10.1086/651009}{What bandwidth do I need for my image?},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{122} 207--214.
\item
Lang,~D., Hogg,~D.~W., Mierle,~K., Blanton,~M., \& Roweis,~S., 2010,
\doi{10.1088/0004-6256/139/5/1782}{\project{Astrometry.net}:\ Blind astrometric calibration of arbitrary astronomical images},
\textit{Astron.\,J.}\ \textbf{139} 1782--1800.
This documents a highly used piece of software, with more than 10,000 downloads; it is obtainable from \url{http://astrometry.net/}.
\item
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011,
\doi{10.1214/10-AOAS439}{Extreme deconvolution:\ Inferring complete distribution functions from noisy, heterogeneous and incomplete observations},
\textit{Ann.\,Appl.\,Stat.}\ \textbf{5} 1657--1677.
\item
Lang,~D. \& Hogg,~D.~W., 2012,
\doi{10.1088/0004-6256/144/2/46}{Searching for comets on the World Wide Web:\ The orbit of 17P/Holmes from the behavior of photographers},
\textit{Astron.\,J.} \textbf{144} 46.
\item
Foreman-Mackey,~D., Hogg,~D.~W., Lang,~D., \& Goodman,~J., 2013,
\doi{10.1086/670067}{\project{emcee}:\ The MCMC Hammer},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{125} 306--312.
This documents a highly used piece of software, currently getting more than ten journal citations per month;
  it is obtainable from \url{http://dan.iel.fm/emcee/current/}.
\item
Brewer,~B.~J., Foreman-Mackey,~D., \& Hogg,~D.~W., 2013,
\doi{10.1088/0004-6256/146/1/7}{Probabilistic catalogs for crowded stellar fields},
\textit{Astron.\,J.} \textbf{146} 7.
\item
Ambikasaran,~S., Foreman-Mackey,~D., Greengard,~L., Hogg,~D.~W., \& O'Neil,~M.,
\arxiv{1403.6015}{Fast direct methods for Gaussian Processes and the analysis of NASA \satellite{Kepler} Mission Data},
\textit{IEEE Transactions on Pattern Analysis and Machine Intelligence}, submitted.
\end{list}

\hoggheading{Other papers and projects}
\begin{list}{}{\hogglist}
\item\label{pub:Hogg2001}
Hogg,~D.~W., Finkbeiner,~D.~P., Schlegel,~D.~J., \& Gunn,~J.~E., 2001,
\doi{10.1086/323103}{A photometricity and extinction monitor at the Apache Point Observatory},
\textit{Astron.\,J.}\ \textbf{122} 2129--2138.
\item
Blanton,~M.~R., Hogg,~D.~W., \etal, 2003,
\doi{10.1086/375528}{The broadband optical properties of galaxies with redshifts $0.02<z<0.2$},
\textit{Astrophys.\,J.}\ \textbf{594} 186--207.
\item\label{pub:Hogg2004}
Hogg,~D.~W. \etal, 2004,
\doi{10.1086/381749}{The dependence on environment of the olor--magnitude relation of galaxies},
\textit{Astrophys.\,J.\,Lett.}\ \textbf{601} L29--L32.
\item\label{pub:Eisenstein2005}
Eisenstein,~D.~J., Zehavi,~I., Hogg,~D.~W., \etal, 2005,
\doi{10.1086/466512}{Detection of the baryon acoustic peak in the large-scale correlation function of \project{Sloan Digital Sky Survey} Luminous Red Galaxies},
\textit{Astrophys.\,J.}\ \textbf{633} 560--574.
\item\label{pub:Masjedi2006}
Masjedi,~M., Hogg,~D.~W., \etal, 2006,
\doi{10.1086/503536}{Very small-scale clustering and merger rate of luminous red galaxies},
\textit{Astrophys.\,J.}\ \textbf{644} 54--60.
\item
Bovy,~J., Hogg,~D.~W., \& Rix,~H.-W., 2009,
\doi{10.1088/0004-637X/704/2/1704}{Galactic masers and the Milky Way circular velocity},
\textit{Astrophys.\,J.}\ \textbf{704} 1704--1709.
\item
Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W., 2010,
\doi{10.1088/0004-637X/712/1/260}{Constraining the Milky Way potential with a 6-D phase-space map of the GD-1 stellar stream},
\textit{Astrophys.\,J.}\ \textbf{712} 260--273.
\end{list}

\hoggheading{Recent collaborators in academia:}\\[0.5ex]
Jo~Bovy (Institute for Advanced Study);
Brendon Brewer (University of Auckland);
Rob~Fergus (New York University);
Marla~Geha (Yale University);
Jonathan~Goodman (New York University);
Leslie~Greengard (New York University);
Joe Hennawi (Max-Planck-Institut f\"ur Astronomie);
Kathryn~Johnston (Columbia University);
Dustin~Lang (Carnegie Mellon University);
Ben~Oppenheimer (American Museum of Natural History);
Hans-Walter~Rix (Max-Planck-Institut f\"ur Astronomie);
David~Schiminovich (Columbia University);
Adi~Zolotov (Hebrew University).

\end{document}
