% This file is part of the OpenK2 project.
% Copyright 2015 David W. Hogg.

% to-do
% -----
% - update to THIS PROJECT; ie, relevant publications

\documentclass[12pt]{article}
\setlength{\headheight}{0ex}
\setlength{\headsep}{0ex}
\input{hogg_nasa}
\pagestyle{empty}
\raggedright

\begin{document}
\noindent
\textbf{David W. Hogg} \hfill http://cosmo.nyu.edu/hogg/

\paragraph{Professional Preparation:}~\\
\begin{tabular}{@{}llll}
\textsl{postdoctoral}
  & Institute for Advanced Study          & Astrophysics & 1997--2001\\
\textsl{graduate}
  & California Institute of Technology    & Physics & PhD 1998\\
\textsl{undergraduate}
  & Massachusetts Institute of Technology & Physics & SB 1992\\
\end{tabular}
           
\paragraph{Appointments:}~\\
\begin{tabular}{@{}lll}
Professor of Physics and Data Science & New York University & 2015--present \\
Professor of Physics                  & New York University & 2014--2015 \\
Associate Professor of Physics        & New York University & 2007--2014 \\
Assistant Professor of Physics        & New York University & 2001--2007 \\
Long-term member in Astrophysics      & Institute for Advanced Study & 1997--2001 \\
\end{tabular}

\paragraph{Relevant Recent Research Efforts:}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
\observatory{Hubble Space Telescope} Archival Resarch grant (AR-13250; Hogg, PI),
\textit{Probabilistic Self-Calibration of the WFC3 IR Channel},
2013--2016.
\item
NASA Astrophysics Data Analysis Grant (NNX12AI50G; Hogg, PI),
\textit{The Lives and Deaths of Planets and Stars in the Value-Added UV Photon Catalog},
2012--2015.
\item
NSF Cyber-Enabled Discovery Type I Grant (IIS-1124794; Hogg, PI),
\textit{A Unified Probabilistic Model of Astronomical Imaging,}
2011--2014.
\end{itemize}

\paragraph{Relevant Recent Publications:}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Lang,~D., Hogg,~D.~W., \& Schlegel,~D.~J., 2016,
{\project{WISE} photometry for 400 million \project{SDSS} sources},
\textit{Astron.\,J.}\ \textbf{151} 36.
\item
Ness,~M., Hogg,~D.~W., \etal, 2015,
{\project{The~Cannon}: A data-driven
approach to stellar label determination},
\textit{Astrophys.\,J.}\ \textbf{808} 16.
\item
Foreman-Mackey,~D. \etal, 2015,
{A systematic search for transiting planets in the \project{K2} data},
\textit{Astrophys.\,J.}\ \textbf{806} 215.
\item
Foreman-Mackey,~D., Hogg,~D.~W., \etal, 2013,
{\project{emcee}:\ The MCMC Hammer},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{125} 306--312.
\item
Hogg,~D.~W., Myers,~A.~D., \& Bovy,~J., 2010,
{Inferring the eccentricity distribution},
\textit{Astrophys.\,J.}\ \textbf{725} 2166--2175.
\item
Lang,~D., Hogg,~D.~W., \etal, 2010,
{\project{Astrometry.net}:\ Blind astrometric calibration of arbitrary astronomical images},
\textit{Astron.\,J.}\ \textbf{139} 1782--1800.
\end{itemize}

\paragraph{Synergistic Activities:}
$\bullet$~Co-author of multiple open-source code bases for astronomical and general data analysis.
$\bullet$~Author of pedagogical texts on arXiv.org for cosmology, fundamental
measurement in astronomy, and statistical inference.
$\bullet$~Co-organizer of Hack Days at the American Astronomical Society winter meetings (2013--present), and
the annual \project{Astro Hack Week} (2014--present).

\paragraph{Graduate Students Advised:}
$\bullet$~Jo Bovy (PhD 2011);
$\bullet$~Daniel Foreman-Mackey (PhD 2015);
$\bullet$~Fengji Hou (PhD 2014);
$\bullet$~Tao Jiang (PhD 2012);
$\bullet$~Dustin Lang (PhD 2009; co-advised by Sam Roweis, Toronto);
$\bullet$~Alex Malz (current);
$\bullet$~Morad Masjedi (PhD 2007);
$\bullet$~Mohammadjavad Vakili (current);
$\bullet$~Dun Wang (current);
$\bullet$~Ronin RongYing Wu (PhD 2011);
$\bullet$~Adi Zolotov (PhD 2011; co-advised by Beth Willman, Haverford).

\paragraph{Postdoctoral Scholars Advised:}
$\bullet$~Andreas Berlind (now at Vanderbilt);
$\bullet$~Michael Blanton (now at NYU);
$\bullet$~Ross Fadely (current);
$\bullet$~Gabriel Perez-Giz (former);
$\bullet$~Erin Sheldon (now at Brookhaven);
$\bullet$~Beth Willman (now at Haverford).

\end{document}
