% This file is part of the OpenK2 project.
% Copyright 2015 David W. Hogg.

% to-do
% -----
% - update to THIS PROJECT; ie, relevant publications

\documentclass[12pt]{article}
\setlength{\headheight}{0ex}
\setlength{\headsep}{0ex}
\input{hogg_nasa}
\pagestyle{empty}
\raggedright

\begin{document}
\noindent
\textbf{David W. Hogg} \hfill http://cosmo.nyu.edu/hogg/
\smallskip

\paragraph{Professional Preparation:}~\\
\begin{tabular}{llll}
\textsl{postdoctoral}
  & Institute for Advanced Study          & Astrophysics & 1997--2001\\
\textsl{graduate}
  & California Institute of Technology    & Physics & PhD 1998\\
\textsl{undergraduate}
  & Massachusetts Institute of Technology & Physics & SB 1992\\
\end{tabular}
           
\paragraph{Appointments:}~\\
\begin{tabular}{lll}
Professor           & Department of Physics, New York University & 2014--present \\
Associate Professor & Department of Physics, New York University & 2007--2014 \\
Assistant Professor & Department of Physics, New York University & 2001--2007 \\
Long-term member    & Astrophysics, Institute for Advanced Study & 1997--2001 \\
\end{tabular}

\paragraph{Administrative Appointments:}~\\
\begin{tabular}{ll}
Deputy Director, NYU Center for Data Science & 2014--present \\
Executive Director, Moore--Sloan Data Science Environment at NYU & 2014--present \\
Director of Undergraduate Studies, NYU Department of Physics & 2008--present \\
\end{tabular}

\paragraph{Relevant Recent Research Efforts:}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
\observatory{Hubble Space Telescope} Archival Resarch grant (AR-13250; Hogg, PI),
\textit{Probabilistic Self-Calibration of the WFC3 IR Channel},
2013--2016.
\item
NASA Astrophysics Data Analysis Grant (NNX12AI50G; Hogg, PI),
\textit{The Lives and Deaths of Planets and Stars in the Value-Added UV Photon Catalog},
2012--2015.
\item
NSF Cyber-Enabled Discovery Type I Grant (IIS-1124794; Hogg, PI),
\textit{A Unified Probabilistic Model of Astronomical Imaging,}
2011--2014.
\item
NSF Astronomy and Astrophysics Research Grant (AST-0908357; Hogg, PI),
\textit{Dynamical models from kinematic data:\ The Milky Way Disk and Halo,}
2009--2011.
\item
NASA Astrophysics Data Analysis Grant (NNX08AJ48G; Hogg, PI),
\textit{Multi-wavelength astrometric catalog built from NASA data,}
2008--2011.
\end{itemize}

\paragraph{Selected Publications:}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Montet,~B.~T., Morton,~T.~D., Foreman-Mackey,~D., Johnson,~J.~A.,
Hogg,~D.~W., Bowler,~B.~P., Latham,~D.~W., Bieryla,~A., Mann,~A.~W., 2015,
{Stellar and planetary properties of \observatory{K2} Campaign 1
  candidates and validation of 18 Systems, including a planet
  receiving earth-like insolation},
\textit{Astrophys.\,J.}\ in press (arXiv:1503.07866).
\item
Foreman-Mackey,~D., Montet,~B.~T., Hogg,~D.~W., Morton,~T.~D., Wang,~D., \& Sch\"olkopf,~B., 2015,
{A systematic search for transiting planets in the \observatory{K2} data},
\textit{Astrophys.\,J.}\ in press (arXiv:1502.04715).
\item
Lang,~D., Hogg,~D.~W., Schlegel,~D.~J., 2015,
{\observatory{WISE} photometry for 400 million \observatory{SDSS} sources},
\textit{Astrophys.\,J.}\ submitted (arXiv:1410.7397).
\item
Foreman-Mackey~D., Hogg,~D.~W., \& Morton,~T.~D., 2014,
{Exoplanet population inference and the abundance of Earth analogs from noisy, incomplete catalogs},
\textit{Astrophys.\,J.}\ \textbf{795} 64.
\item
Brewer,~B.~J., Foreman-Mackey,~D., \& Hogg,~D.~W., 2013,
{Probabilistic catalogs for crowded stellar fields},
\textit{Astron.\,J.} \textbf{146} 7.
\item
Foreman-Mackey,~D., Hogg,~D.~W., Lang,~D., \& Goodman,~J., 2013,
{\project{emcee}:\ The MCMC Hammer},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{125} 306--312.
\item
Fadely,~R., Hogg,~D.~W., \& Willman,~B., 2012,
{Star--galaxy classification in multi-band optical imaging},
\textit{Astrophys.\,J.}\ \textbf{760} 15.
\item
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011,
{Extreme deconvolution:\ Inferring complete distribution functions from noisy, heterogeneous and incomplete observations},
\textit{Ann.\,Appl.\,Stat.}\ \textbf{5} 1657--1677.
\item
Bovy,~J., Hennawi,~J.~F., Hogg,~D.~W., \etal, 2011,
{Think outside the color-box:\ Probabilistic target selection and the \project{SDSS-XDQSO} quasar targeting catalog},
\textit{Astrophys.\,J.}\ \textbf{729} 141.
\item
Hogg,~D.~W., Myers,~A.~D., \& Bovy,~J., 2010,
{Inferring the eccentricity distribution},
\textit{Astrophys.\,J.}\ \textbf{725} 2166--2175.
\item
Lang,~D., Hogg,~D.~W., Mierle,~K., Blanton,~M., \& Roweis,~S., 2010,
{\project{Astrometry.net}:\ Blind astrometric calibration of arbitrary astronomical images},
\textit{Astron.\,J.}\ \textbf{139} 1782--1800.
\item
Marshall,~P.~J., Hogg,~D.~W., Moustakas,~L.~A., Fassnacht,~C.~D.,
Brada\v{c},~M., Schrabback,~T., \& Blandford,~R.~D., 2009,
{Automated detection of galaxy-scale gravitational lenses in high-resolution imaging data},
\textit{Astrophys.\,J.}\ \textbf{694} 924--942.
\end{itemize}

\paragraph{Synergistic Activities:}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Co-author of open-source code
for performing tuned Monte Carlo sampling in inference contexts (http://dan.iel.fm/emcee/),
for fast Gaussian process regression and marginalization with enormous matrices (http://dan.iel.fm/george/),
for estimating densities given heterogeneous observations (http://code.google.com/p/extreme-deconvolution/),
for performing classification and target selection (http://code.google.com/p/xdtarget/),
and
for calibration of astronomical images (http://Astrometry.net/).
\item
Author of pedagogical texts on arXiv.org for cosmology, fundamental
measurement in astronomy, and---recently---statistical inference.
These texts are associated with publicly released code.
\item
Co-organizer of Hack Days at the American Astronomical Society winter meetings (2013--present).
Co-organizer of annual \project{Astro Hack Week} (2014--present).
\item
Recipient of various teaching awards: 
NYU ``Golden Dozen'' Teaching Award (2004);
Princeton University Engineering Council Teaching Award (2000);
Caltech Undergraduate Teaching Award (1996).
Gives occasional public lectures to increase public understanding of
science, recently in downtown Abu Dhabi (audience of 250) and at
the NYC ``Secret Science Club'' (audience of 600).
\item
Participant in committees in service to the astronomical community,
recently including
White House OSTP Astronomy and Astrophysics Advisory Committee (2014--present),
\Spitzer\ Science Center Oversight Committee, (2008--present),
\project{Kepler} Exoplanet Occurence Rate review (2014),
and
\project{NASA Extragalactic Database} Users Committee (2006--2013).
\end{itemize}

\paragraph{Graduate Students Advised:}
$\bullet$~Jo Bovy (PhD 2011);
$\bullet$~Daniel Foreman-Mackey (PhD 2015);
$\bullet$~Fengji Hou (PhD 2014);
$\bullet$~Tao Jiang (PhD 2012);
$\bullet$~Dustin Lang (PhD 2009; co-advised by Sam Roweis, Toronto);
$\bullet$~Alex Malz (current);
$\bullet$~Morad Masjedi (PhD 2007);
$\bullet$~Mohammadjavad Vakili (current);
$\bullet$~Dun Wang (current);
$\bullet$~Ronin RongYing Wu (PhD 2011);
$\bullet$~Adi Zolotov (PhD 2011; co-advised by Beth Willman, Haverford).

\paragraph{Postdoctoral Scholars Advised:}
$\bullet$~Andreas Berlind (now at Vanderbilt);
$\bullet$~Michael Blanton (now at NYU);
$\bullet$~Ross Fadely (current);
$\bullet$~Gabriel Perez-Giz (former);
$\bullet$~Erin Sheldon (now at Brookhaven);
$\bullet$~Beth Willman (now at Haverford).

\end{document}
