\documentclass[10pt]{article}
\jot=8pt
\textheight 9.0in
\topmargin -.75in
\oddsidemargin 0.2in
\evensidemargin \oddsidemargin
\pagestyle{empty}
\textwidth 6.2in
\setlength{\parindent}{0pt}  

\setlength{\parskip}{\baselineskip}
\newcommand{\name}{\par\centerline{\large\bf{
Jonathan B. Goodman}}\vspace{.25\baselineskip} \par}
\newcommand{\heading}[1]{\par\noindent{\normalsize\bf{#1}}\par}
%\newcommand{\subheading}[1]{\par{\normalsize{#1}}\par}
\newenvironment{items}[4]%
{\begin{minipage}[t]{#1}
 \begin{list}{ }{ \setlength{\leftmargin}{#2}
                 \setlength{\labelwidth}{#3}
                 \setlength{\labelsep}{#4}
                 \setlength{\parsep}{0in}
                 \setlength{\topsep}{0in}
                 \setlength{\partopsep}{0in}
                 \setlength{\itemsep}{0pt}
                 \setlength{\rightmargin}{0in} } }%
{\end{list}
 \end{minipage}}
\begin{document}
\pagestyle{empty}
\newlength{\noyear}
\settowidth{\noyear}{1984}
\newlength{\nyd}
\settowidth{\nyd}{--1984}
\newlength{\boxw}
\setlength{\boxw}{0.99\textwidth}
\newlength{\boxp}
\setlength{\boxp}{0.9\textwidth}

\name
\vspace{.2in}
\vspace*{-.15in}
\begin{tabbing}
Connecticutx, \=Connecticutx, \=Connecticutx, \=Connecticutx,
\=Connecticutx, \=Connecticutx, \=Connecticutx, \kill

Courant Institute of Mathematical Sciences  \>\>\>\> Tel: (212) 998-3326 \\
New York University  \>\>\>\>  Fax: (212) 995-4121 \\
251 Mercer Street \>\>\>\> Email: goodman@cims.nyu.edu \\
New York, NY 10012 \\
\end{tabbing}

\vspace*{.1in}
%\vfill
\vspace*{.0in}

\heading{Education}
\vspace*{-.1in}
\begin{items}{\boxw}{1.0in}{1.0in}{0.2in}
\item[PhD 1982] Mathematics; 
		{Stanford University}, Stanford, CA		\\
		Advisor:  Andrew Majda
\item[\,\,\,\,BS 1977]  Mathematics;
		{Massachusetts Institute of Technology}, Cambridge, MA	
\end{items}

\vfill
\heading{Positions}
\vspace*{-.1in}
\begin{items}{\boxw}{1.0in}{1.0in}{0.2in}
\item[1990--date\,] Professor of Mathematics;
		New York University,  New York, NY\\
		Chair, Math Finance MS program (1998-2003;2006-2009)

\item[1986--1990] Associate Professor of Mathematics; 
		New York University,  New York, NY	
\item[1984--1986] Assistant Professor of Mathematics; 
		New York University,  New York, NY	
\item[1984] Consultant ; AT\& T Bell Laboratories (summer)
\item[1982--1984] Courant Institute Instructor of Mathematics; 
		New York University,  New York, NY	\\
\item[1978--1982] Teaching Assistant; Stanford University
\item[1979] Numerical Analysis Consultant; SLAC (Stanford U.)
\item[1978--1979] Research Assistant, Computer Science Dept., Stanford U.
\item[1977--1978] MIT Energy Laboratory 
\item[1976] MIT Electronics Systems Laboratory 
\end{items}

\vfill

\heading{Honors}
\vspace*{-.1in}
\begin{items}{\boxw}{1.0in}{1.0in}{0.2in}
\item[1986] Monroe Martin Prize for best paper by recent Ph.D. in Applied
            Mathematics
\item[1986] Alfred P. Sloan Fellowship
\item[1986] NSF Presidential Young Investigator Award
\end{items}

\vfill

\heading{Editorial Boards}
\vspace*{-.1in}
\begin{items}{\boxw}{1.0in}{1.0in}{0.2in}
\item[1988--date]  Communications in Pure and Applied Mathematics
\item[2003--date]  Journal of Computational Finance
\item[1988--1990]  AMS Series on Contemporary Mathematics
\item[1990--1992]  SIAM Journal on Numerical Analysis
\item[1997--1999]  Journal of Computational Physics
\end{items}

\vfill

\heading{Ph.D. students}
\vspace*{-.1in}
Kalman Meth (1987), Steven Cohn (1989), Gabriella Puppo (1990), 
Shlomo Engleberg(1994), Peter Schultz (1996), Hyeonggi Lee (1996),
Juan Carlos Aguilar (1999), Savio Rodriges (1999),  Kazuhiro Iwasawa (2000),
Ramon Plaza (2002), Jose Antonio Perez
(2004), Sivan Rottenstreich (2005), Ed Hayes (2007), Kranthi Gade (2008),
Ethan Atkins (2010), Sangmin Lee (2010), Lin Li (2011), 
Jose Vidal Alcala (2011), SungJae Jun(2011), Yuan Yuan (2012), 
Bo Zhu (2013), Fengji Hou (2014), 
Hongliang Liu (expected 2015), Mike Lewis (expected 2015),
Hongbin Park (expected 2015), Yair Daon (expected 2016)


\vfill

\heading{MS students}
\vspace*{-.1in}
24 students to date

\newpage
\heading{Recent collaborators}
\vspace*{-.1in}


John Bell, Marcus Day, Daniel Foreman-Mackey, Pierre Germain, Ray Grout, Salman Habib,
David Hogg, Fengji Hou, Phillipe LeFloch, Kevin Lin, Matthias Morzfeld, Daniel Ostrov, 
Christian Schwab, Jonathan Weare


\vspace*{.1in}
\heading{Publications (a selection)}
\vspace*{-.1in}
\begin{enumerate}

\item
``Convergence of the Random Vortex Method", Comm. Pure Appl. Math.
40 (1987) pp. 189 - 220

\item
``Multi Grid Monte Carlo III. Two dimensional O(4)-Symmetric Nonlinear 
sigma-Model"
(with Robert Edwards, Sabino Ferreira, and Alan Sokal), 
Nuclear Physics B (Proc. Suppl.)
\item
``Stability of the Kuramoto Sivashinsky and related systems",  
Comm. Pure. Appl. Math. 47(1994) pp. 293-306

\item
``Modeling pinchoff and reconnection in a Hele-Shaw cell II: analysis
and simulation in the nonlinear regime'', (with Hyeong-Gi Lee and John 
Lowengrub) Physics of Fluids 14(2), Feb., 2002.

\item
``An efficient interpolation
algorithm on anisotropic grids for functions with jump continuities in
2-D", (with Juan Carlos Aguilar), Applied Numerical Mathematics 55 (2), Oct.
2005.

\item
``Coupling control variates for MCMC", (with Kevin Lin). J. Comput. Phys.
228 (19), 2009.
\item
``Balancing small transaction costs with loss of optimal allocation
in single and multiple stock portfolios", (with Dan Ostrov). 
SIAM J. Applied Math., 70(6), pp. 1977-1998, 2010.
\item
``Ensemble samplers with affine invariance", (with Jonathan Weare).
Comm. Applied Math Comp. Sci., 5(1), pp. 65--80, 2010.
\item
``An Affine-Invariant Sampler for Exoplanet Fitting and Discovery in
Radial Velocity Data", (with Fengji Hou, David W. Hogg, Jonathan Weare,
Christian Schwab).  arXiv:1104.2612v2. The Astrophysical J. 745:198, 2012.

\item
``Small-noise analysis and symmetrization of implicit Monte Carlo
samplers", witth Kevin K. LIn and Matthias Morzfeld.  Submitted.


\end{enumerate}

\end{document}
