\documentclass[letterpaper,12pt]{article}
\setlength{\headheight}{0ex}
\setlength{\headsep}{0ex}
\input{hogg_nasa}
\pagestyle{empty}
\begin{document}

\paragraph{David W. Hogg} http://cosmo.nyu.edu/hogg/

\paragraph{Professional Preparation:}~\\[3pt]
\begin{tabular}{llll}
\textsl{postdoctoral}
  & Institute for Advanced Study          & Astrophysics & 1997--2001\\
\textsl{graduate}
  & California Institute of Technology    & Physics & PhD 1998\\
\textsl{undergraduate}
  & Massachusetts Institute of Technology & Physics & SB 1992\\
\end{tabular}
           
\paragraph{Appointments:}~\\[3pt]
\begin{tabular}{lll}
Professor           & Department of Physics, New York University & 2014--present \\
Associate Professor & Department of Physics, New York University & 2007--2014 \\
Assistant Professor & Department of Physics, New York University & 2001--2007 \\
Long-term member    & Astrophysics, Institute for Advanced Study & 1997--2001 \\
\end{tabular}

\paragraph{Administrative Roles:}~\\[3pt]
\begin{tabular}{ll}
Deputy Director, NYU Center for Data Science & 2014--present \\
Executive Director, Moore--Sloan Data Science Environment at NYU & 2014--present \\
Director of Undergraduate Studies, NYU Department of Physics & 2008--present \\
\end{tabular}

\paragraph{Relevant Publications}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Bovy,~J. \etal, 2012,
{The Milky Way's circular-velocity curve between 4 and 14~kpc from \project{APOGEE} data},
\textit{Astrophys.\,J.}\ \textbf{759} 131.
\item
Yoon,~J.~H., Johnston,~K.~V., \& Hogg,~D.~W., 2011,
{Clumpy streams from clumpy halos:\ Detecting missing satellites with cold stellar structures}, 
\textit{Astrophys.\,J.}\ \textbf{731} 58.
\item
Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W., 2010,
{Constraining the Milky Way potential with a 6-D phase-space map of the GD-1 stellar stream},
\textit{Astrophys.\,J.}\ \textbf{712} 260--273.
\item
Bell,~E.~F., \etal, 2008,
{The accretion origin of the Milky Way's stellar halo},
\textit{Astrophys.\,J.}\ \textbf{680} 295--311.
\item
Willman,~B., Blanton,~M.~R., West,~A.~A., Dalcanton,~J.~J, Hogg,~D.~W., Schneider,~D.~P., Wherry,~N., Yanny,~B., \& Brinkmann,~J., 2005,
{A new Milky Way companion:\ Unusual globular cluster or extreme dwarf satellite?},
\textit{Astron.\,J.}\ \textbf{129} 2692--2700.
\item
Foreman-Mackey,~D., Hogg,~D.~W., Lang,~D., \& Goodman,~J., 2013,
{\project{emcee}:\ The MCMC Hammer},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{125} 306--312.
\item
Bovy,~J., Rix,~H.-W., \& Hogg,~D.~W., 2012,
{The Milky Way has no distinct thick disk},
\textit{Astrophys.\,J.}\ \textbf{751} 131.
\item
Lang,~D., Hogg,~D.~W., Mierle,~K., Blanton,~M., \& Roweis,~S., 2010,
{\project{Astrometry.net}:\ Blind astrometric calibration of arbitrary astronomical images},
\textit{Astron.\,J.}\ \textbf{139} 1782--1800.
\item
Masjedi,~M., Hogg,~D.~W., \etal, 2006,
{Very small-scale clustering and merger rate of luminous red galaxies},
\textit{Astrophys.\,J.}\ \textbf{644} 54--60.
\item
Eisenstein,~D.~J., Zehavi,~I., Hogg,~D.~W., \etal, 2005,
Detection of the baryon acoustic peak in the large-scale correlation
function of Sloan Digital Sky Survey Luminous Red Galaxies,
\textit{Astrophys.\,J.}\ \textbf{633} 560--574.
\end{itemize}

\paragraph{Synergistic Activities}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Co-wrote and currently run \project{Astrometry.net} to calibrate
astronomical images; currently have more than one thousand users and
an open-source code base.
\item
Co-author of open-source code for estimating densities
given heterogeneous observations
(http://code.google.com/p/extreme-deconvolution/), for performing
classification and target selection
http://code.google.com/p/xdtarget/, and for performing very fast Monte
Carlo sampling in inference contexts (\textit{arXiv:}1104.2612).
\item
Author of pedagogical texts on arXiv.org for cosmology, fundamental
measurement in astronomy, and---recently---statistical inference.
These texts are associated with publicly released code.
\item
Organizer of Hack Days at the American Astronomical Society winter meetings (2013--present).
Co-organizer of annual Astro Hack Week (2014--present).
\item
Recipient of various teaching awards: 
NYU ``Golden Dozen'' Teaching Award (2004);
Princeton University Engineering Council Teaching Award (2000);
Caltech Undergraduate Teaching Award (1996).
Gives occasional public lectures to increase public understanding of
science, most recently in downtown Abu Dhabi (audience of 250) and at
the NYC ``Secret Science Club'' (audience of 600).
\item
Participant in committees in service to the astronomical community:
White House OSTP Astronomy and Astrophysics Advisory Committee (2014--present);
\Spitzer\ Science Center Oversight Committee, (2008--present); 
\project{Kepler} Exoplanet Occurence Rate review (2014);
\project{NASA Extragalactic Database} Users Committee (2006--2013);
Panel Chair, \Spitzer\ Time Allocation Committee (2005).
\end{itemize}

\paragraph{Recent Collaborators and Co-editors:}
$\bullet$~Ruth~Angus (Oxford);
$\bullet$~Tom~Barclay (Ames);
$\bullet$~Brendon~Brewer (Auckland);
$\bullet$~Kevin~Bundy (IPMU);
$\bullet$~Richard~Cool (MMT);
$\bullet$~Rob~Fergus (NYU);
$\bullet$~Julianne~Dalcanton (UW);
$\bullet$~Rebekah~Dawson (Berkeley);
$\bullet$~Roberto~Decarli (MPIA);
$\bullet$~Morgan~Fouesneau (UW);
$\bullet$~Dustin~Lang (CMU);
$\bullet$~Phil~Marshall (KIPAC);
$\bullet$~Benjamin~Montet (Caltech);
$\bullet$~Benjamin~Oppenheimer (AMNH);
$\bullet$~Hans-Walter~Rix (MPIA);
$\bullet$~David~Schiminovich (Columbia);
$\bullet$~Michael~Schneider (LLNL);
$\bullet$~Bernhard~Sch\"olkopf (MPI-IS);
$\bullet$~Daniel~Weisz (UCSC);
$\bullet$~Kenneth~Wong (Arizona);
$\bullet$~Ann~Zabludoff (Arizona).

\paragraph{Graduate Advisor and Postdoctoral Sponsor:}
$\bullet$~Roger~Blandford (Caltech PhD advisor; now at Stanford);
$\bullet$~John~Bahcall (IAS postdoc sponsor; now deceased).

\paragraph{Graduate Students Advised:}
$\bullet$~Jo~Bovy (PhD 2011);
$\bullet$~Daniel~Foreman-Mackey (current student);
$\bullet$~Fengji~Hou (PhD 2014);
$\bullet$~Tao~Jiang (PhD 2012);
$\bullet$~Dustin~Lang (PhD 2009; co-advised by Sam Roweis, Toronto);
$\bullet$~Morad~Masjedi (PhD 2007);
$\bullet$~Ronin~RongYing Wu (PhD 2011);
$\bullet$~Adi~Zolotov (PhD 2011; co-advised by Beth Willman, Haverford).

\paragraph{Postdoctoral Scholars Advised:}
$\bullet$~Andreas~Berlind (now at Vanderbilt);
$\bullet$~Michael~Blanton (now at NYU);
$\bullet$~Ross~Fadely (current);
$\bullet$~Gabriel~Perez-Giz (currently);
$\bullet$~Erin~Sheldon (now at Brookhaven);
$\bullet$~Beth~Willman (now at Haverford).

\end{document}
