\documentclass[letterpaper,12pt]{article}
\setlength{\headheight}{0ex}
\setlength{\headsep}{0ex}
\input{hogg_nasa}
\pagestyle{empty}
\begin{document}

\paragraph{David W. Hogg} http://cosmo.nyu.edu/hogg/

\paragraph{Professional Preparation:}~\\[3pt]
\begin{tabular}{llll}
\textsl{postdoctoral}
  & Institute for Advanced Study          & Astrophysics & 1997--2001\\
\textsl{graduate}
  & California Institute of Technology    & Physics & PhD 1998\\
\textsl{undergraduate}
  & Massachusetts Institute of Technology & Physics & SB 1992\\
\end{tabular}
           
\paragraph{Appointments:}~\\[3pt]
\begin{tabular}{lll}
Professor           & Department of Physics, New York University & 2014--present \\
Associate Professor & Department of Physics, New York University & 2007--2014 \\
Assistant Professor & Department of Physics, New York University & 2001--2007 \\
Long-term member    & Astrophysics, Institute for Advanced Study & 1997--2001 \\
\end{tabular}

\paragraph{Administrative Roles:}~\\[3pt]
\begin{tabular}{ll}
Deputy Director, NYU Center for Data Science & 2014--present \\
Executive Director, Moore--Sloan Data Science Environment at NYU & 2014--present \\
Director of Undergraduate Studies, NYU Department of Physics & 2008--present \\
\end{tabular}

\paragraph{Ten Recent, Relevant Publications}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Foreman-Mackey~D., Hogg,~D.~W., \& Morton,~T.~D., 2014,
{Exoplanet population inference and the abundance of Earth analogs from noisy, incomplete catalogs},
\textit{Astrophys.\,J.}\ \textbf{795} 64.
\item
Fergus,~R., Hogg,~D.~W., Oppenheimer,~R., Brenner,~D. \& Pueyo,~L., 2014,
{\project{S4}: A spatial-spectral model for speckle suppression},
\textit{Astrophys.\,J.}\ \textbf{794} 161.
\item
Price-Whelan,~A.~M., Hogg,~D.~W., Johnston,~K.~V., \& Hendel,~D., 2014,
{Inferring the gravitational potential of the Milky Way with a few precisely measured stars},
\textit{Astrophys.\,J.}\ \textbf{794} 4.
\item
Brewer,~B.~J., Foreman-Mackey,~D., \& Hogg,~D.~W., 2013,
{Probabilistic catalogs for crowded stellar fields},
\textit{Astron.\,J.} \textbf{146} 7.
\item
Foreman-Mackey,~D., Hogg,~D.~W., Lang,~D., \& Goodman,~J., 2013,
{\project{emcee}:\ The MCMC Hammer},
\textit{Pubs.\,Astr.\,Soc.\,Pac.}\ \textbf{125} 306--312.
\item
Fadely,~R., Hogg,~D.~W., \& Willman,~B., 2012,
{Star--galaxy classification in multi-band optical imaging},
\textit{Astrophys.\,J.}\ \textbf{760} 15.
\item
Bovy,~J. \etal, 2012,
{Photometric redshifts and quasar probabilities from a single, data-driven generative model},
\textit{Astrophys.\,J.}\ \textbf{749} 41.
\item
Hou,~F., Goodman,~J., Hogg,~D.~W., Weare,~J., \& Schwab,~C., 2012,
{An affine-invariant sampler for exoplanet fitting and discovery in radial velocity data},
\textit{Astrophys.\,J.}\ \textbf{745} 198.
\item
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011,
{Extreme deconvolution:\ Inferring complete distribution functions from noisy, heterogeneous and incomplete observations},
\textit{Ann.\,Appl.\,Stat.}\ \textbf{5} 1657--1677.
\item
Hogg,~D.~W., Myers,~A.~D., \& Bovy,~J., 2010,
{Inferring the eccentricity distribution},
\textit{Astrophys.\,J.}\ \textbf{725} 2166--2175.
\end{itemize}

\paragraph{Synergistic Activities}
\begin{itemize}\setlength{\itemsep}{0pt}
\item
Co-author of open-source code
for performing tuned Monte Carlo sampling in inference contexts (http://dan.iel.fm/emcee/),
for fast Gaussian process regression and marginalization with enormous matrices (http://dan.iel.fm/george/),
for estimating densities given heterogeneous observations (http://code.google.com/p/extreme-deconvolution/),
for performing classification and target selection (http://code.google.com/p/xdtarget/),
and
for calibration of astronomical images (http://Astrometry.net/).
\item
Author of pedagogical texts on arXiv.org for statistical inference.
These texts are associated with publicly released code.
\item
Organizer of Hack Days at the American Astronomical Society winter meetings (2013--present).
Co-organizer of annual Astro Hack Week (2014--present).
\item
Recipient of NYU ``Golden Dozen'' Teaching Award (2004);
Princeton University Engineering Council Teaching Award (2000);
Caltech Undergraduate Teaching Award (1996).
Recent public lectures in downtown Abu Dhabi (audience of 250) and at
the NYC ``Secret Science Club'' (audience of 600).
\item
Participant in committees in service to the astronomical community,
recently including
White House OSTP Astronomy and Astrophysics Advisory Committee (2014--present),
\Spitzer\ Science Center Oversight Committee, (2008--present),
\project{Kepler} Exoplanet Occurence Rate review (2014),
\project{NASA Extragalactic Database} Users Committee (2006--2013),
and
Panel Chair, \Spitzer\ Time Allocation Committee (2005).
\end{itemize}

\paragraph{Recent Collaborators and Co-editors:}
$\bullet$~Ruth Angus (Oxford);
$\bullet$~Tom Barclay (Ames);
$\bullet$~Brendon Brewer (Auckland);
$\bullet$~Kevin Bundy (IPMU);
$\bullet$~Charlie Conroy (Harvard);
$\bullet$~Julianne Dalcanton (UW);
$\bullet$~Rebekah Dawson (Berkeley);
$\bullet$~Roberto Decarli (MPIA);
$\bullet$~Rob Fergus (NYU);
$\bullet$~Morgan Fouesneau (UW);
$\bullet$~Marla Geha (Yale);
$\bullet$~Jonathan Goodman (NYU);
$\bullet$~Joseph P. Hennawi (MPIA);
$\bullet$~Kathryn Johnston (Columbia);
$\bullet$~Andreas Kuepper (Columbia);
$\bullet$~Dustin Lang (CMU);
$\bullet$~Khee-Gan Lee (MPIA);
$\bullet$~Elisabeta Lusso (Arcetri);
$\bullet$~Phil Marshall (KIPAC);
$\bullet$~Timothy Morton (Princeton);
$\bullet$~Melissa Ness (MPIA);
$\bullet$~Michael O'Neil (NYU);
$\bullet$~Rebecca Oppenheimer (AMNH);
$\bullet$~Hans-Walter Rix (MPIA);
$\bullet$~Robyn W. Sanderson (Columbia);
$\bullet$~David Schiminovich (Columbia);
$\bullet$~David J. Schlegel (LBL);
$\bullet$~Michael Schneider (LLNL);
$\bullet$~Bernhard Sch\"olkopf (MPI-IS);
$\bullet$~Daniel Weisz (UCSC);
$\bullet$~Kenneth Wong (Arizona);
$\bullet$~Ann Zabludoff (Arizona).

\paragraph{Graduate Advisor and Postdoctoral Sponsor:}
$\bullet$~Roger Blandford (Caltech PhD advisor; now at Stanford);
$\bullet$~John Bahcall (IAS postdoc sponsor; now deceased).

\paragraph{Graduate Students Advised:}
$\bullet$~Jo Bovy (PhD 2011);
$\bullet$~Daniel Foreman-Mackey (current);
$\bullet$~Fengji Hou (PhD 2014);
$\bullet$~Tao Jiang (PhD 2012);
$\bullet$~Dustin Lang (PhD 2009; co-advised by Sam Roweis, Toronto);
$\bullet$~Alex Malz (current);
$\bullet$~Morad Masjedi (PhD 2007);
$\bullet$~Mohammadjavad Vakili (current);
$\bullet$~Dun Wang (current);
$\bullet$~Ronin RongYing Wu (PhD 2011);
$\bullet$~Adi Zolotov (PhD 2011; co-advised by Beth Willman, Haverford).

\paragraph{Postdoctoral Scholars Advised:}
$\bullet$~Andreas Berlind (now at Vanderbilt);
$\bullet$~Michael Blanton (now at NYU);
$\bullet$~Ross Fadely (current);
$\bullet$~Gabriel Perez-Giz (current);
$\bullet$~Erin Sheldon (now at Brookhaven);
$\bullet$~Beth Willman (now at Haverford).

\paragraph{Current and Pending Support:}
$\bullet$~CDI-Type I: A Unified Probabilistic Model of Astronomical Imaging (PI; Co-PI Rob Fergus, NYU),
NSF, 675,000\,USD, 2011--2015, 0.33 month;
$\bullet$~The Lives and Deaths of Planets and Stars in the Value-Added UV Photon Catalog,
NASA, 473,705\,USD, 2012--2015, 0.33 month;
$\bullet$~Probabilistic Self-Calibration of the WFC3 IR Channel,
STScI, 119,988\,USD, 2013--2016, 0.33 month;
$\bullet$~Moore--Sloan Data Science Environment at NYU,
Moore Foundation and Sloan Foundation, 12,500,000\,USD, 2013-2018, 2.0 month.

\end{document}
