\documentclass[letterpaper,12pt]{article}
\usepackage{url}
\setlength{\headheight}{0ex}
\setlength{\headsep}{0ex}
\input{hogg_nasa}
\pagestyle{empty}
\begin{document}

\begin{center}
    {\bf \large Michael O'Neil}\\
    {\em Biosketch}
\end{center}

\bigskip

\noindent
Courant Institute of Mathematical Sciences\\
New York University\\
251 Mercer St., \#1122\\
New York, NY 10012\\
(212) 998-3125\\
\url{oneil@cims.nyu.edu}\\
\url{www.cims.nyu.edu/~oneil}\\

\bigskip

\noindent
{\bf Professional preparation}
\begin{tabbing}
Cornell University \qquad \= Mathematics \qquad \qquad \qquad \= A.B. 
    \qquad \=2003\\
Yale University \> Applied Mathematics \> M.Phil. \> 2005 \\
Yale University \> Applied Mathematics \> Ph.D. \> 2007 \\
\end{tabbing}



\noindent
{\bf Appointments}
\begin{tabbing}
2014 - \qquad \qquad \= {\em Assistant Professor of Mathematics} \\
\> Courant Institute \& Polytechnic School of Engineering, New York
University\\
 2012 - 2014 \> {\em Courant Instructor}\\
  \>Courant Institute, New York University \\
2010 - 2012 \> {\em Associate Research Scientist} \\
       \> Courant Institute, New York University\\
2007 - 2010 \> {\em Assistant trader and quantitative
            researcher}\\
    \> Susquehanna International Group, LLP \\
    \> Bala Cynwyd, PA 19004\\
\end{tabbing}




\noindent
{\bf Relevant Publications}
\begin{enumerate}
    \item {\em Fast Direct Methods for Gaussian Processes and the
        Analysis of NASA Kepler Mission Data},\\
            S. Ambikasaran, D. Foreman-Mackey,
                L. Greengard, D. W. Hogg, and M. O'Neil,\\
                submitted (under review),
            \url{http://arxiv.org/abs/1403.6015}.
    \item {\em Fast symmetric factorization of hierarchical matrices
            with applications},\\
            S. Ambikasaran and M. O'Neil,\\
            submitted (under review), \url{http://arxiv.org/abs/1405.0223}. 
    \item {\em An algorithm for the rapid evaluation of special 
            function transforms},\\
            M. O'Neil, F. Woolfe, and V. Rokhlin, \\
            Appl. Comput. Harmon. Anal. 28(2): 203-226, 2010.
    \item {\em Debye sources and the numerical solution of the time 
        harmonic Maxwell equations, II},\\
        C. L. Epstein,  L. Greengard, and M. O'Neil,\\ 
        Comm. Pure Appl. Math. 66(5):753-789, 2013.
    \item {\em Quadrature by Expansion: A New Method for the Evaluation 
        of Layer Potentials},
        A. Kl\"ockner, A. Barnett, L. Greengard, and M. O'Neil),\\
        J. Comput. Phys. 252:332-349, 2013.
    \item {\em A generalized Debye source approach to electromagnetic
        scattering in layered media},\\
        M. O'Neil,\\
        J. Math. Phys. 55, 012901, 2014.\\
\end{enumerate}




   
\noindent
{\bf Collaborators} \par \noindent
Sivaram~Ambikasaran (NYU),
Alexander~Barnett (Dartmouth),
Antoine~Cerfon (NYU),
Charles~L.~Epstein (UPenn),
Daniel~Foreman-Mackey (NYU),
Zydrunas~Gimbutas (NIST),
Leslie~Greengard (NYU),
David~W.~Hogg (NYU),
Andreas~Kl\"ockner (UIUC),
Gunnar~Martinsson (UC Boulder),
Vladimir~Rokhlin (Yale),
Jon~Wilkening (Berkeley)\\


\noindent
{\bf Graduate and Postdoctoral Advisors}\par \noindent
Vladimir Rokhlin (Ph.D. advisor, Yale University)\\
Leslie Greengard (Postdoctoral advisor, New York University)\\

\noindent
{\bf Current Ph.D. and Postgraduate Sponsorship} \par \noindent
Sivaram Ambikasaran, (Postdoctoral, NYU)\\
Sunli Tang (Ph.D., NYU)\\

\end{document}
