\documentclass[12pt, letterpaper]{article}
\usepackage{fancyhdr, graphicx}
\setlength{\headsep}{0ex}
\input{hogg_nasa}
  \renewcommand{\headrulewidth}{0pt}
  \pagestyle{empty}

\begin{document}

\paragraph{Data management plan:}
This project will not produce new data \foreign{per se}, but it does produce
methods, code, and [HOGG INSERT THINGS HERE].
All of the code, documentation, data, and all written documents relating to the
project will be maintained in publicly visible version-control repositories on
the GitHub and GitLab platforms. These repositories can be cloned and reproduced by any
user in the world.

The code and documents will be licensed for re-use by others under permissive
licenses (MIT for code, CC-by for documents). This permits re-use, reproducibility,
and re-publication by others.

Although we do not expect to deliver radial-velocity measurements directly, if
we do create any catalog of stellar radial velocities, we will return these to
the \HARPS\ Archive (or whatever is the relevant hosting archive). [HOGG DOES THIS MAKE SENSE??]
We will also publish them in GitHub repositories, and license them for re-use.

Any cite-able data, code, or documentation will be delivered to Zenodo for
preservation and creation of a cite-able DOI.
All documents and papers will be submitted to the arXiv and also \AAS-family
journals for long-term preservation, cite-ability, and publication.

The PI has a strong tradition in his group for open science, performing
all data analysis and writing in the open, with a strong preference for working
in open and public data sets. He also supports maintainance and development of
open-source software systems of wide use in the astronomical community.

\end{document}
