\documentclass[12pt]{article}
\setlength{\headheight}{0ex} % must be *before* \input{hogg_nasa}
\setlength{\headsep}{0ex} % must be *before* \input{hogg_nasa}
\input{hogg_nasa}
\begin{document}\sloppy\sloppypar\thispagestyle{empty}

\noindent
\textbf{Facilities and Resources}
\smallskip

The NYU Department of Physics is part of the NYU Faculty of Arts \&
Science and includes the Center for Cosmology and Particle Physics,
the Center for Soft Matter Research, and the Center for Quantum Phenomena.
A wide variety of sponsored
research activities take place in the Department.  These range from
the theoretical to the pragmatic, and include a broad spectrum of
interactions with such disciplines as biology, medicine, chemistry,
applied mathematics, and computer science.

In addition to benefitting from this research activity directly and
indirectly, as a member of the Department, the PI receives through the
Department staff support for clerical work, post-award grant support,
and for computing (expanded upon below).

\paragraph{Computing Equipment:}

The astrophysics group at NYU maintains some high-performance computers
for the PI, and substantial storage machines.
The latter contain more than 100~Tb of disk space, most of which is filled
with astrophysical imaging data.  The Physics Department maintains a
state-of-the-art controlled-environment computer room that houses and
protects the computers that will be used in this project.

\paragraph{General computer Resources:}

The University has hired a Director of Scientific Computing for the
NYU Center for Cosmology and Particle Physics (the PI's home).  His
responsibilities include management of the cluster and the data
servers, and overall management and supervision of the Center's
computer system.  He will oversee the computer hardware used in this
project.

A variety of computing resources are available within the Physics
Department, including UNIX workstations, laptops, laser color
printers, etc.  The Department runs a network of several hundred
desktop workstations, file servers, in addition to the multi-processor
computational server mentioned above.

Computational needs are also supported through the University's
Academic Computing Services, a unit of the NYU-wide Information
Technology Services offering an additional wide range of computational
resources in support of research and instruction.  These include a
variety of computing platforms, including several high-performance
multi-CPU systems, and scientific software.  Consultants are available
to assist in the use of these resources.

Finally, the NYU Center for Data Science (where the PI has an appointment)
maintains some specialized high-performance machines, including a GPU
cluster for machine-learning
applications. These resources could be relevant to the proposed project,
especially as some development might be optimized for GPUs.

\paragraph{Office Space:}

The offices of the PI and GRA will be located in the NYU Physics
Department.

\paragraph{Library Resources:}

In addition to an enormous book collection, the NYU libraries hold
current subscriptions to hundreds of hardcopy and electronic journals.
Libraries provides access to all journals relevant to this project and such
databases as MathSciNet, the Web of Science (science citation index),
and the ACM Digital Library. They also provide resources to support
researchers with preservation of data, code, and models.

\paragraph{Meeting spaces:}

The Center for Cosmology and Particle Physics, the NYU Libraries, the
NYU Center for Data Science, and the NYU Kimmel Center for Student
Life all have spaces that are available to the PI for meetings and
workshops.

\paragraph{Experimental and Hardware Facilities:}

The Department has large experimental facilities, including machining,
imaging, and clean facilities, but none of these are directly relevant
to this project, except insofar as they are part of the rich
intellectual and research atmosphere.

\end{document}
