% This file is part of the GetRichOrDieTrying project.
% Copyright 2014 David W. Hogg (NYU) and MJ Vakili (NYU).

\documentclass[12pt]{article}
\setlength{\headheight}{0ex} % must be *before* \input{hogg_nsf}
\setlength{\headsep}{0ex} % must be *before* \input{hogg_nsf}
\input{hogg_nsf}
\begin{document}\sloppy\sloppypar\thispagestyle{empty}

\noindent
\textsf{David W. Hogg / \textbf{New Probabilistic Methods for Observational Cosmology}} % must be synchronized with description.tex
\smallskip

As new cosmological surveys get underway,
we expect to measure smaller signals with larger numbers of galaxies
observed (individually) at lower confidence.
These goals require us to perform data
analyses that are as information-preserving as possible.
This proposal is to create new methods for cosmological data analysis
that will permit the cosmology community to perform cosmological
inferences using not lossy, derived data products (N-sigma galaxy
catalogs, best-fit redshifts, correlation function point estimates),
but something much closer to the original imaging and spectroscopic
data.
The new methods will have to be informed by principles of
probabilistic inference, and also good applied-mathematics technology.

There are three (related) toolsets comprising this proposal.
\package{Toolset~1} is for reconstruction and marginalization
of cosmological density fields, given typical data.
The density field could be the mass, galaxy, or 
neutral-gas density field, or the two-dimensional projected mass density,
depending on the tracer.
This Toolset requires repeated solves and determinants for
million-element dense linear systems.  For the first time, we
have applied-math tools for this.

\package{Toolset~2} is for cosmological inference that make proper
use of probabilistic information about galaxy and quasar redshifts.
All justified methods involve ``forward modeling'' (which, in turn,
capitalizes on \package{Toolset~1}).  The tools will also, simultaneously,
\emph{improve} probabilistic redshift information, and provide
informative imputation of missing redshifts, thereby producing
predictions and tools for smaller-scale scientific questions.

\package{Toolset~3} is for propagation of probabilistic image-level quantities---%
galaxy shapes and the point-spread function (PSF)---%
into weak-lensing studies of large-scale structure.
The tools will permit inference from survey data conditional priors over
galaxy shapes and PSFs, and posteriors relevant every object.
They will also make use of these in a justified forward-modeling
measurement of the shear field and cosmological parameters.

\smallskip\noindent
\textbf{$\bullet$~Intellectual Merit:}
Projects in Stage-III (\eg, \boss, \des)
and Stage-IV (\eg, \lsst, \euclid, \wfirst)
are starting to see the limits of point estimates and rigid catalogs;
they are considering probabilistic output for redshifts,
shapes, and photometry.
The toolsets from this project will be among the first practical methods for
cosmological inference and large-scale structure measurement that can
make full and proper (justified) use of these probabilistic outputs.
For the first time, it will be possible to perform
simultaneous inference or refinement of catalog-level properties along with
large-scale structure and cosmological inferences.
Simultaneous inference is expected to significantly \emph{reduce statistical
biases in cosmological measurements},
and also reduce variance in catalog-level quantities.

Full probabilistic inference at the survey level has so far appeared
computationally intractable.
The methods developed here will make use of---and propagate more
generally in the field---collaborations with applied mathematics that
make the impossible possible.

\smallskip\noindent
\textbf{$\bullet$~Broader Impacts:}
The toolsets will be papers and methods but also open-source codebases
that will benefit all cosmological projects, and also projects that
are not in cosmology at all.
The development of the toolsets will aid in the development of
standards for generating and delivering probabilistic outputs, which
will influence Stage-IV projects.

The PI will continue his practice of producing pedagogical papers on
inference, data analysis, and computational statistics in the physical
sciences.
These reach a range of populations inside and outside academia.
He will also continue to work with his students to organize and staff
influential data-analysis Hack Days at scientific meetings.

\end{document}
